### 2019-04-20 paper/mybib.tex

@article{cui2016continuous,
  title={Continuous online sequence learning with an unsupervised neural network model},
  author={Cui, Yuwei and Ahmad, Subutai and Hawkins, Jeff},
  journal={Neural computation},
  volume={28},
  number={11},
  pages={2474--2504},
  year={2016},
  publisher={MIT Press}
}

@article{waibel1995phoneme,
  title={Phoneme recognition using time-delay neural networks},
  author={Waibel, Alexander and Hanazawa, Toshiyuki and Hinton, Geoffrey and Shikano, Kiyohiro and Lang, Kevin J},
  journal={Backpropagation: Theory, Architectures and Applications},
  pages={35--61},
  year={1995},
  publisher={Lawrence Erlbaum Assoc. Mahwah, NJ}
}

@article{diehl2015unsupervised,
  title={Unsupervised learning of digit recognition using spike-timing-dependent plasticity},
  author={Diehl, Peter U and Cook, Matthew},
  journal={Frontiers in computational neuroscience},
  volume={9},
  pages={99},
  year={2015},
  publisher={Frontiers}
}

@book{gerstner2002spiking,
  title={Spiking neuron models: Single neurons, populations, plasticity},
  author={Gerstner, Wulfram and Kistler, Werner M},
  year={2002},
  publisher={Cambridge university press}
}

@article{cruz2012energy,
  title={Energy-efficient neuron, synapse and STDP integrated circuits},
  author={Cruz-Albrecht, Jose M and Yung, Michael W and Srinivasa, Narayan},
  journal={IEEE transactions on biomedical circuits and systems},
  volume={6},
  number={3},
  pages={246--256},
  year={2012},
  publisher={IEEE}
}

@article{maass1997networks,
  title={Networks of spiking neurons: the third generation of neural network models},
  author={Maass, Wolfgang},
  journal={Neural networks},
  volume={10},
  number={9},
  pages={1659--1671},
  year={1997},
  publisher={Elsevier}
}

@inproceedings{mikolov2013distributed,
  title={Distributed representations of words and phrases and their compositionality},
  author={Mikolov, Tomas and Sutskever, Ilya and Chen, Kai and Corrado, Greg S and Dean, Jeff},
  booktitle={Advances in neural information processing systems},
  pages={3111--3119},
  year={2013}
}

@article{hochreiter1997long,
  title={Long short-term memory},
  author={Hochreiter, Sepp and Schmidhuber, J{\"u}rgen},
  journal={Neural computation},
  volume={9},
  number={8},
  pages={1735--1780},
  year={1997},
  publisher={MIT Press}
}

@article{mozer1995focused,
  title={A focused backpropagation algorithm for temporal},
  author={Mozer, Michael C},
  journal={Backpropagation: Theory, architectures, and applications},
  volume={137},
  year={1995}
}

@incollection{rojas1996backpropagation,
  title={The backpropagation algorithm},
  author={Rojas, Raul},
  booktitle={Neural networks},
  pages={149--182},
  year={1996},
  publisher={Springer}
}

@inproceedings{diehl2015fast,
  title={Fast-classifying, high-accuracy spiking deep networks through weight and threshold balancing},
  author={Diehl, Peter U and Neil, Daniel and Binas, Jonathan and Cook, Matthew and Liu, Shih-Chii and Pfeiffer, Michael},
  booktitle={2015 International Joint Conference on Neural Networks (IJCNN)},
  pages={1--8},
  year={2015},
  organization={IEEE}
}

@article{beliaev2007time,
  title={Time series prediction using chaotic neural networks on the CATS benchmark},
  author={Beliaev, Igor and Kozma, Robert},
  journal={Neurocomputing},
  volume={70},
  number={13-15},
  pages={2426--2439},
  year={2007},
  publisher={Elsevier}
}